Marlowe: financial contracts on blockchain\thanks{This work is part of the Cardano project and is supported by IOHK,~\url{https://iohk.io}}

### 2018-07-26 tex/paper.tex
Added single type decl for names
Fix some typos

The outer primitive \haskellinline{When} waits until the condition -- its first argument -- 
becomes true; in this case, the condition is the that either two participants choose \haskellinline{refund} or two 
participants choose \haskellinline{pay}. The second argument of the \haskellinline{When} is itself another 
\haskellinline{Contract}, which is performed after the condition of the \haskellinline{When} has been met, and it makes 
the payment if two participants chose \haskellinline{pay}, otherwise it redeems previous money commitments.

@@

On the other hand, at each step there are -- potentially, at least -- a variety of inputs available from the 
participants themselves. These include commitments of currency (or ``cash''), redemption of commitments, and claims of 
payments by a participant. Moreover it is also possible for a participant to input an arbitrary value (which we term a 
``choice''). The particular inputs at a given step are described by a value of type \haskellinline{Input}.

@@

In this section, we show how Marlowe can be used for designing a contract that incentivises saving. We use the 
following example: \haskellinline{bob} offers 20 ADA to \haskellinline{alice} in exchange for saving 100 ADA for 100 
blocks. \haskellinline{alice} will be able to withdraw the savings after block 20, but if that option is chosen, then 
\haskellinline{bob} will not pay the extra 20 ADA to participant \haskellinline{alice}.

@@

In our implementation we make things more specific: the money paid for the item is 450 ADA, and it must be committed by 
\haskellinline{alice} before block \haskellinline{10}; it will be refunded to participant \haskellinline{alice} if 
there is no consensus before block \haskellinline{90}.

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
An identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
And we can create identifiers for all the participants:
\begin{haskellcode}
alice :: Person
alice = 1

bob :: Person
bob = 2

carol :: Person
carol = 3
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} 
to be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either \haskellinline{alice} deserves a refund, which 
we represent with the number $0$; or \haskellinline{bob} deserves a payment, which we represent with the 
number $1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as 
their participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then, we can easily write a function that returns an observation that is \haskellinline{True} if and only if at least 
one of the participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{val} as 
follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is \haskellinline{True} if and only if 
at least two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed 
in choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose alice bob carol 0)
                                 (two_chose alice bob carol 1))
                          90
                          (Choice (two_chose alice bob carol 1)
                                  (Pay iP1 alice bob (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

@@

For Marlowe to be usable in practice, users need to be able to understand how contracts will behave once deployed to 
the blockchain, but without doing the deployment. We can do that by simulating their behaviour off-chain, interactively 
stepping through the evaluation of a contract in a browser. We do this in two stages, first transforming an embedded 
contact (using features of Haskell) to a pure Marlowe contract, and then interactively stepping through that contract.

@@

If we have contracts that involve, for example, the oil price then we may need to convert a published price in USD into 
ADA, using a value for the prevailing USD/ADA exchange rate. This will require a computation: in this case, a 
multiplication. If Marlowe is to be stand-alone then we need to be extend it with arithmetic and other operations, but 
we expect Marlowe to be embedded in a suitable language that provides these facilities.

>>>>

The outer primitive \haskellinline{When} waits until the condition -- its first argument -- 
becomes true; in this case, the condition is that either two participants choose \haskellinline{refund} or two 
participants choose \haskellinline{pay}. The second argument of the \haskellinline{When} is itself another 
\haskellinline{Contract}, which is performed after the condition of the \haskellinline{When} has been met, and it makes 
the payment if two participants chose \haskellinline{pay}, otherwise it redeems previous money commitments.

@@

On the other hand, at each step there are -- potentially, at least -- a variety of inputs available from the 
participants themselves. These include commitments of currency (or ``cash''), redemption of commitments, and claims of 
payments by a participant. Moreover, it is also possible for a participant to input an arbitrary value (which we term a 
``choice''). The particular inputs at a given step are described by a value of type \haskellinline{Input}.

@@

In this section, we show how Marlowe can be used for designing a contract that incentivises saving. We use the 
following example: \haskellinline{bob} offers 20 ADA to \haskellinline{alice} in exchange for saving 100 ADA for 100 
blocks. \haskellinline{alice} will be able to withdraw the savings after block 20, but if that option is chosen, then 
\haskellinline{bob} will not pay the extra 20 ADA to \haskellinline{alice}.

@@

In our implementation we make things more specific: the money paid for the item is 450 ADA, and it must be committed by 
\haskellinline{alice} before block \haskellinline{10}; it will be refunded to \haskellinline{alice} if 
there is no consensus before block \haskellinline{90}.

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
An identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
And we can create identifiers for all the participants:
\begin{haskellcode}
alice, bob, carol :: Person
alice = 1
bob   = 2
carol = 3
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} 
to be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either \haskellinline{alice} deserves a refund, which 
we represent with the number $0$; or \haskellinline{bob} deserves a payment, which we represent with the 
number $1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as 
their participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then, we can easily write a function that returns an observation that is \haskellinline{True} if and only if at least 
one of the participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{val} as 
follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is \haskellinline{True} if and only if 
at least two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed 
in choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose alice bob carol 0)
                                 (two_chose alice bob carol 1))
                          90
                          (Choice (two_chose alice bob carol 1)
                                  (Pay iP1 alice bob (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

@@

For Marlowe to be usable in practice, users need to be able to understand how contracts will behave once deployed to 
the blockchain, but without doing the deployment. We can do that by simulating their behaviour off-chain, interactively 
stepping through the evaluation of a contract in a browser. We do this in two stages, first transforming an embedded 
contract (using features of Haskell) to a pure Marlowe contract, and then interactively stepping through that contract.

@@

If we have contracts that involve, for example, the oil price then we may need to convert a published price in USD into 
ADA, using a value for the prevailing USD/ADA exchange rate. This will require a computation: in this case, a 
multiplication. If Marlowe is to be stand-alone then we need to extend it with arithmetic and other operations, but 
we expect Marlowe to be embedded in a suitable language that provides these facilities.
